---
title: "R Notebook"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r}
knitr::opts_chunk$set(cache = T)
library(SIS)
library(DT)
library(stringr)
library(parallel)
library(CoxHD)
library(mg14)
ncore <- detectCores()
```
# Data Preparation

Load data

```{r}
dataFrame <- readRDS('dataFrame.rds')
os  <- readRDS('os.rds')
load('AMLSG_Clinical_Anon.RData')
```

REMOVE NA

```{r}

poorMansImpute <- function(x) {x[is.na(x)] <- mean(x, na.rm=TRUE); return(x)}
dataFrame <- as.data.frame(sapply(dataFrame, poorMansImpute))

```

Cluster

```{r}

library(dplyr)
gene_data=as.matrix(dataFrame)
d1=dist(gene_data)
hc1=hclust(d1,"ward.D")
clusters <- cutree(hc1,k=3)
table(clusters)
dataFrame <- cbind(dataFrame,clusters)

```


Strip out OS and form X

```{r}
dim(dataFrame)
str(dataFrame$TPL_os)
str_subset(names(dataFrame),"^TPL")


datX<- as.matrix(subset(dataFrame,select = !grepl("^TPL",names(dataFrame))))
datX.df <- subset(dataFrame,select = !grepl("^TPL",names(dataFrame)))
## str_subset(names(trainX),"^TPL")



```

Prepare Survival Data

```{r}
## From 1.3.3
## os

t <- clinicalData$Time_Diag_TPL
t[is.na(t) | !clinicalData$TPL_Phase %in% "CR1" | !clinicalData$TPL_type %in% c("ALLO","FREMD") ] <- Inf ## Only allografts in CR1
o <- clinicalData$OS
tplIndexOs <-  t < o
osTD <-  Surv(time = rep(0, nrow(clinicalData)), time2=pmin(o, t), event=ifelse(tplIndexOs, 0, clinicalData$Status) )
osTD <- rbind(osTD, 
        Surv(time=t[which(tplIndexOs)],
                time2=o[which(tplIndexOs)], 
                event=clinicalData$Status[which(tplIndexOs)])
)
osTD = Surv(osTD[,1],osTD[,2],osTD[,3])
rm(o,t)
tplSplitOs <- c(1:nrow(clinicalData), which(tplIndexOs))
osYr <- os
osYr[,1] <- osYr[,1]/365
osYrTD <- osTD
osYrTD[,1] <- osYrTD[,1]/365

```

# CoXHD::CPSS

## Read Group


```{r}
dataList <- readRDS("dataList.rds")
groups <- unlist(sapply(names(dataList), function(x) rep(x, ncol(dataList[[x]]))))
groups[grepl("^(t_)|(inv)", colnames(dataFrame)) &! grepl(":", colnames(dataFrame))] <- "Fusions"
groups[groups=="Cytogenetics"] <- "CNA"
groups <- factor(groups)
names(groups) <- colnames(dataFrame)
table(groups)

## 2.3.1
dataFrameOsTD <- dataFrame[tplSplitOs,]
dataFrameOsTD[which(tplIndexOs),grep("TPL", colnames(dataFrameOsTD), value=TRUE)] <- 0
mainGroups <- grep("[A-Z][a-z]+[A-Z]",levels(groups), invert=TRUE, value=TRUE)
mainGroups

mainIdx <- groups %in% mainGroups
osIdx <- !grepl("TPL", colnames(dataFrame)) ## Exclude TPL from OS analyses..
whichRFXOs <- which((colSums(dataFrame)>=8 | mainIdx) & osIdx) # ie, > 0.5%
mainIdxOs <- mainIdx & osIdx
osTDIdx <- !grepl("TPL_efs", colnames(dataFrame))
whichRFXOsTD <- which((colSums(dataFrame)>=8 | mainIdx) & osTDIdx) # ie, > 0.5%
mainIdxOsTD <- mainIdx & osTDIdx
whichRFXOsGG <- which((colSums(dataFrame)>=8 | mainIdxOs) & osIdx & groups %in% c(mainGroups,"GeneGene")) # ie, > 0.5%

```

Prepare for multistage model

```{r}
## 3.5.1
library(msSurv)
## Package dependency: graph
## http://www.bioconductor.org/packages/release/bioc/html/graph.html


d <- sapply(1:nrow(clinicalData), function(i){
            i <<- i
            t <- c(as.numeric(clinicalData[i,c("CR_date","Recurrence_date","Date_LF")]) - as.numeric(clinicalData$ERDate[i]))
            o <- order(t, na.last=NA)
            stages <- c(1:3,0)
            r <- stages[c(1, o+1)]
            if(clinicalData$Status[i])
                r[length(r)] <- r[length(r)-1] +3
            tt <- c(0,t[o])
            if(length(o)==0)
                return(c(rep(NA,7),i))
            s <- cbind(id=i, stop=tt[-1], start.stage=r[-length(r)], end.stage=r[-1])[diff(tt)!=0,]
            #s <- cbind(time1 = tt[-length(tt)], time2=tt[-1], death=c(rep(0, length(o)-1), clinicalData$Status[i]), outer(0:(length(o)-1), r[-3], `>=`)+0, i=i)[diff(tt)!=0,]
            return(s)
        })
d <- as.data.frame(do.call("rbind",d))
nodes <- as.character(1:6)
edges <- list(`1`=list(edges=c("2","4")), `2`=list(edges=c("3","5")), `3`=list(edges="6"), `4`=list(edges=NULL), `5`=list(edges=NULL),`6`=list(edges=NULL))
struct <-  new("graphNEL", nodes = nodes, edgeL = edges, edgemode = "directed")
msurv <- msSurv(d, struct, bs = FALSE)

## Include gene:gene interaction terms
##
whichRFXOsTDGG <- which((colSums(dataFrame)>=8 | mainIdxOsTD) & osTDIdx & groups %in% c(mainGroups,"GeneGene")) # ie, > 0.5%
coxRFXFitOsTDGGc <- CoxRFX(dataFrameOsTD[,whichRFXOsTDGG], osTD, groups[whichRFXOsTDGG], which.mu=mainGroups) ## allow only the main groups to have mean different from zero.. 

## Times fro allografts 3.5.2

alloIdx <- clinicalData$TPL_type %in% c("ALLO","FREMD") # only allografts
alloTimeCR1 <- clinicalData$Time_1CR_TPL + .5 # +.5 to make > 0
alloTimeCR1[!alloIdx | (clinicalData$TPL_date < clinicalData$Recurrence_date & !clinicalData$TPL_Phase %in% c("CR1","RD"))] <- NA

whichRFXRel <- whichRFXOsTDGG[grep("TPL",names(whichRFXOsTDGG), invert=TRUE)] #mainIdx & !grepl("TPL", names(dataFrame)) & groups!="Nuisance"
t <- clinicalData$Recurrence_date
t[is.na(t)] <- as.Date(1e6, origin="2000-01-01")
relData <- MakeTimeDependent(dataFrame[whichRFXRel], timeEvent=alloTimeCR1, timeStop=as.numeric(pmin(t, clinicalData$Date_LF) - clinicalData$CR_date), status=!is.na(clinicalData$Recurrence_date)+0)
relData$transplantCR1 <- relData$event
relData$event <- NULL
relData$transplantRel <- 0
nrdData <- MakeTimeDependent(dataFrame[whichRFXRel], timeEvent=alloTimeCR1, timeStop=as.numeric(pmin(t, clinicalData$Date_LF) - clinicalData$CR_date), status=is.na(clinicalData$Recurrence_date) & clinicalData$Status)
nrdData$transplantCR1 <- nrdData$event
nrdData$event <- NULL
nrdData$transplantRel <- 0
alloTimeRel <- clinicalData$TPL_date - clinicalData$Recurrence_date + .5 # +.5 to make > 0
alloTimeRel[!alloIdx | (clinicalData$TPL_date < clinicalData$Recurrence_date & !clinicalData$TPL_Phase %in% c("CR1","RD"))] <- NA
i <- !is.na(clinicalData$Recurrence_date)
prdData <- MakeTimeDependent(dataFrame[i,whichRFXRel], timeEvent=alloTimeRel[i], timeStop=as.numeric(clinicalData$Date_LF- clinicalData$Recurrence_date)[i], status=clinicalData$Status[i])
prdData$transplantCR1 <- rep(0,nrow(prdData))
w <- sub("\\.1","",rownames(relData))[relData$status==1 & relData$transplantCR1==1]
prdData$transplantCR1[sub("\\.1","",rownames(prdData)) %in% w] <- 1
prdData$transplantRel <- prdData$event
prdData$event <- NULL
w <- which(prdData$time1 == prdData$time2) ## 5 cases with LF=Rec
prdData$time2[w] <- prdData$time2[w] + .5
prdData$time0 <- as.numeric(clinicalData$Recurrence_date-clinicalData$CR_date)[prdData$index]

```

MultiStage Model Prediction Fn: 3.5.5

```{r}
library(Rcpp)

MultiRFX5 <- function(coxRFXNcdTD, coxRFXCrTD, coxRFXNrdTD, coxRFXRelTD, coxRFXPrdTD, data, x =365, tdPrmBaseline = rep(1, ceiling(max(x))+1), tdOsBaseline = rep(1, ceiling(max(x))+1), ciType="analytical"){
    cppFunction('NumericVector computeHierarchicalSurvival(NumericVector x, NumericVector diffS0, NumericVector S1Static, NumericVector haz1TimeDep) {
                    int xLen = x.size();
                    double h;
                    NumericVector overallSurvival(xLen);
                    for(int i = 0; i < xLen; ++i) overallSurvival[i] = 1;
                    for(int j = 1; j < xLen; ++j){
                    h = haz1TimeDep[j-1];
                    for(int i = j; i < xLen; ++i){
                    overallSurvival[i] += diffS0[j-1] * (1-pow(S1Static[i-j], h));
                    }
                    }
                    return overallSurvival;
                    }')
    
    
    
    ## Step 1: Compute KM survival curves and log hazard
    getS <- function(coxRFX, data, max.x=5000) {        
        if(!is.null(coxRFX$na.action)) coxRFX$Z <- coxRFX$Z[-coxRFX$na.action,]
        data <- as.matrix(data[,match(colnames(coxRFX$Z),colnames(data)), drop=FALSE])
        r <- PredictRiskMissing(coxRFX, data, var="var2")
        H0 <- basehaz(coxRFX, centered = FALSE)
        hazardDist <- splinefun(H0$time, H0$hazard, method="monoH.FC")
        x <- c(0:ceiling(max.x))
        S <- exp(-hazardDist(x))
        return(list(S=S, r=r, x=x, hazardDist=hazardDist, r0 = coxRFX$means %*% coef(coxRFX)))
    }
    kmCr <- getS(coxRFX = coxRFXCrTD, data = data, max.x=max(x))
    kmEs <- getS(coxRFX = coxRFXNcdTD, data = data, max.x=max(x))
    kmCir <- getS(coxRFX = coxRFXRelTD, data = data, max.x=max(x))
    kmNrm <- getS(coxRFX = coxRFXNrdTD, data = data, max.x=max(x))
    kmPrs <- getS(coxRFX = coxRFXPrdTD, data = data, max.x=max(x))
    
    xx <- 0:ceiling(max(x))
    
    sapply(1:nrow(data), function(i){
                ## Step 2: Adjust curves for competing risks, accounting for hazard
                crAbs <-  cumsum(c(1,diff(kmCr$S^exp(kmCr$r[i,1]))) * kmEs$S ^ exp(kmEs$r[i,1]))
                esAbs  <- cumsum(c(1,diff(kmEs$S^exp(kmEs$r[i,1]))) * kmCr$S ^ exp(kmCr$r[i,1])) ## array times x nrow(data)
                cirCrAbs <- cumsum(c(1,diff(kmCir$S^exp(kmCir$r[i,1]))) * kmNrm$S ^ exp(kmNrm$r[i,1]))
                nrsCrAbs <- cumsum(c(1,diff(kmNrm$S^exp(kmNrm$r[i,1]))) * kmCir$S ^ exp(kmCir$r[i,1])) ## array times x nrow(data)
                
                ## Step 3: Compute hierarchical survival
                ### Prs         
                rsCrAbs <- computeHierarchicalSurvival(x = xx, diffS0 = diff(cirCrAbs), S1Static = kmPrs$S, haz1TimeDep = tdPrmBaseline * exp(kmPrs$r[i,1]))
                
                ## Confidence intervals (loglog)
                PlogP2 <- function(x) {(x * log(x))^2}
                errOs <- kmNrm$r[i,2] * PlogP2(kmNrm$S^exp(kmNrm$r[i,1])) * (1-(1-kmCir$S ^ exp(kmCir$r[i,1]))) * (1-kmPrs$S ^ exp(kmPrs$r[i,1]))^2 + kmCir$r[i,2] * PlogP2(kmCir$S ^ exp(kmCir$r[i,1])) * (1-kmPrs$S ^ exp(kmPrs$r[i,1]))^2 * (kmNrm$S ^ exp(kmNrm$r[i,1]))^2 +  kmPrs$r[i,2] * PlogP2(kmPrs$S ^ exp(kmPrs$r[i,1])) * (1-kmCir$S ^ exp(kmCir$r[i,1]))^2 * (kmNrm$S ^ exp(kmNrm$r[i,1]))^2 
                sdOsCr <- sqrt(errOs / PlogP2(1-(1-nrsCrAbs)-(1-rsCrAbs)))
                
                
                ### Overall survival from enrollment
                nrsEr <- computeHierarchicalSurvival(x = xx, diffS0 = diff(crAbs), S1Static = nrsCrAbs, haz1TimeDep = tdOsBaseline)
                rsEr <- computeHierarchicalSurvival(x = xx, diffS0 = diff(crAbs), S1Static = rsCrAbs, haz1TimeDep = tdOsBaseline)
                cirEr <- computeHierarchicalSurvival(x = xx, diffS0 = diff(crAbs), S1Static = cirCrAbs, haz1TimeDep = tdOsBaseline)
                cbind(deathInErFromEr=1-esAbs, deathInCrFromEr=1-nrsEr, deathInRelFromEr=1-rsEr, aliveInRelFromEr=1-cirEr-(1-rsEr), aliveInCrFromEr=1-crAbs - (1-cirEr) - (1-nrsEr),
                        deathInCrFromCr = 1-nrsCrAbs, deathInRelapseFromCr=(1-rsCrAbs), aliveInRelapseFromCr = (1-cirCrAbs) - (1-rsCrAbs), osInCrFromCrSd=sdOsCr
                )
            }, simplify='array')
}

```





```{r}



# Repeated Splitting

```{r,eval=F}
datX <- subset(dataFrameOsTD,select=grepl("^TPL",names(dataFrameOsTD)))
## Repeat 10 times
b <- 10
prderCPSS <- prderSIS <- matrix(0,nrow=b,ncol=1)
AUCCPSS <- AUCSIS <- matrix(0,nrow=b,ncol=1)

for ( i in 1:b)
{
    
    ## Split into training testing dataset 30% training and 30% testing
    idx <- sample(c(1:nrow(dataFrameOsTD),size=0.8*nrow(dataFrameOsTD)))
    traindat <- dataFrameOsTD[idx,]
    trainX <- 
    testdat <- dataFrameOsTD[-idx,]


    require(CoxHD)
    require(mg14)
    scope <-  c("Genetics","CNA","Treatment","Fusions")
    CPSSfit <- CoxCPSSInteractions(X=trainX.df,surv=os.train,scope=which(groups %in% scope),mc.cores=ncore)
    CPSSresult <- names(CPSSfit$coxph$coefficients)

    ## Model Performance

### AUC

    require(survAUC)
    ## SIS formula

    sisfit <- SIS(x=trainX,y=os.train,family='cox',penalty='lasso',tune='bic',iter=FALSE)
    SISresult <- names(trainX.df)[sisfit$ix]


    SISVar <- paste(SISresult,collapse='+')
    SISForm <- as.formula(paste("os.train ~",SISVar))

    ## CPSS formula
    CPSSVar <- str_split(CPSSfit$coxph$formula,"~")[[3]]
    CPSSForm <- as.formula(paste("os.train ~",CPSSVar))

    ## SIS
    times <- seq(10,1825,1) ## times according to author

    tr.fit.sis <- coxph(SISForm, data=trainX.df,x=T,y=T)
    lp.sis <- predict(tr.fit.sis)
    lpnew.sis <- predict(tr.fit.sis, newdata=testX.df)
    surv.rsp.sis <- os.train
    surv.rsp.new.sis <- os.test


    AUC.sis <- AUC.uno(surv.rsp.sis,surv.rsp.new.sis, scale(lpnew.sis),times)$auc

    ## CPSS
    tr.fit.cpss <- coxph(CPSSForm, data=trainX.df,x=T,y=T)
    lp.cpss <- predict(tr.fit.cpss)
    lpnew.cpss <- predict(tr.fit.cpss, newdata=testX.df)
    surv.rsp.cpss <- os.train
    surv.rsp.new.cpss <- os.test


    AUC.cpss <- AUC.uno(surv.rsp.cpss,surv.rsp.new.cpss, scale(lpnew.cpss),times)$auc

    ##plot(x=times,y=AUC.cpss,type='l',xlab="Days",ylab="AUC",ylim=c(0.4,0.7))
    ##lines(x=times,y=AUC.sis,col=2)
    ##legend("bottomright",c("CPSS","SIS"),lty=1,col=c(1,2))

    prderCPSS[i] <- mean(predErr(surv.rsp.cpss,surv.rsp.new.cpss,lp=lp.cpss, lpnew=lpnew.cpss,times=times)$error)
    prderSIS[i] <- mean(predErr(surv.rsp.sis,surv.rsp.new.sis,lp=lp.sis, lpnew=lpnew.sis,times=times)$error)
    AUCCPSS[i] <- mean(AUC.cpss)
    AUCSIS[i] <- mean(AUC.sis)

}


```

```{r}
## Load pre-run data
err <- readRDS("err.rds")

boxplot(err,main="Prediction Error")

```

```{r}
auc.prd <- readRDS('aucprd.rds')
boxplot(auc.prd,main="AUC")
```
